---
title: "Survival Analysis for seed germination"
author: "Ruth Delgado-Dávila"
output: html_document
---

# Inbreeding Depression

This analysis is part of the assessment of temporal variation in inbreeding depression in *Ipomoea hederacea* (see [Delgado-Dávila and Martén-Rodríguez, 2021](https://www.researchgate.net/publication/356285929_A_test_of_the_reproductive_assurance_hypothesis_in_Ipomoea_hederacea_does_inbreeding_depression_counteract_the_benefits_of_self-pollination) for details).

* Plant population: Catemaco, Veracruz, Mexico - 18°24′N, 95°06′W, 340 m a.s.l.  

* Study years: 2012-2013  

```{r warning=FALSE, include=FALSE}
library(tidyverse)
library(survival) # Surv function
library(flexsurvcure) # flexsurvcure function
library(cmprsk) # cuminc function
```

## Data

```{r echo=TRUE, message=FALSE, warning=FALSE}
germination <- read.csv("https://zenodo.org/record/5091713/files/Germination_ID.csv?download=1")
germination$Year <- as.factor(germination$Year)
head(germination)
tail(germination)
```

### Selecting viable seeds 

Next analysis only include viable seeds (seed weight > 12 g, see [Delgado-Dávila *et al.* 2016](https://www.researchgate.net/publication/308203245_Variation_in_floral_morphology_and_plant_reproductive_success_in_four_Ipomoea_species_Convolvulaceae_with_contrasting_breeding_systems)).  


```{r echo=TRUE, message=FALSE, warning=FALSE}
# Viable Seeds
viableSeeds <- germination %>% filter(S_Weight >= 11)
summary(viableSeeds$S_Weight)
```

### Plotting germinated seeds 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Germinated Seeds Subset
germinated <- filter(viableSeeds, Germination == 1)
mycolors = viridis::cividis(2, begin = 0.3, end = 0.7)
# plot median, minimum, and maximum of germinated seeds per year
plot <- germinated %>% 
  ggplot(aes(x = Treatment, y = G_Day, color = Treatment, fill = Treatment))
plot <- plot + stat_summary(fun.ymin = min, 
                            fun.ymax = max, 
                            fun.y = "median", 
                            shape = 21, 
                            size = 1) 
plot <- plot + facet_wrap(~ Year)
plot <- plot + scale_color_manual(values = mycolors) 
plot <- plot + scale_fill_manual(values = mycolors) 
plot <- plot + coord_flip()
plot <- plot + theme_light()
plot <- plot + labs(
  x = "pollination treatment", 
  y = "days"
)
plot
```

## Cure Model 

A parametric mixture cure model is applied to compare probability survival curves per treatment by year.
The analysis required the creation of survival objects, which are created using the ```Surv``` function. A survival object consists of two columns:  

* First column is the time or **censored time**.

* Second column is the **censoring indicator**, indicating right censored data.  

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Cure model for year 2012
g12 <- dplyr::filter(viableSeeds, Year == "2012")
Surv_Cure12 <- flexsurvcure(
  Surv(
    G_Day, Germination
    ) ~ Treatment, 
  data = g12, 
  link="logistic", 
  dist="lnorm", 
  mixture = T
  )
# plot(Surv_Cure12)
# Cure model for year 2013
g13 <- dplyr::filter(viableSeeds, Year == "2013")
Surv_Cure13 <- flexsurvcure(
  Surv(
    G_Day, Germination
    ) ~ Treatment, 
  data = g13, 
  link="logistic", 
  dist="lnorm", 
  mixture = T
  )
# plot(Surv_Cure13)
```


## Plotting cumulative incidence by treatment 

Cumulative Probability: shows the probability of germination at a given time, and it is defined as:  

$$
\begin{aligned}
F(t) = 1 - S(t)
\end{aligned}
$$
Where $S(t)$ is the survival function, indicating the probability of survive (not germinating) at a given time. 

Following code lines show how to prepare seed data to plot the cumulative probability of germination.  

```{r echo=TRUE, message=FALSE, warning=FALSE}
NewData <- viableSeeds %>% 
  group_by(Year, Treatment) %>% 
  mutate(Total = n()) %>%
  ungroup()
# Number of germinated seeds by treatment
Count <- NewData %>% 
  group_by(Year, Treatment, Total, G_Day) %>% 
  summarise(sum(Germination)) %>%
  ungroup()

colnames(Count) <- c("Year", "Treatment", "Total", "Day", "Germ_seeds")
head(Count)
# Cumulative germination of seeds by treatment in 2012
ACUM <- Count %>% 
  group_by(Year, Treatment) %>% 
  mutate(ACUM = cumsum(Germ_seeds)) %>% 
  mutate(EVENT = ACUM/Total, KM = 1 - EVENT) %>%
  ungroup()
head(ACUM)
```


```{r echo=TRUE, message=FALSE, warning=FALSE}
# Plotting cumulative probability per year
ACUM_12 <- ACUM %>% dplyr::filter(Year == "2012")
plot(ACUM_12$Day, ACUM_12$EVENT, 
     pch = c(19, 1)[as.numeric(as.factor(ACUM_12$Treatment))], 
     cex = 1, 
     axes = F,
     xlim = c(0,122), 
     ylim = c(0,1), 
     ylab = "", 
     xlab = "")
par(mgp = c(axis.title.position = 1, axis.label.position = 0.8, axis.line.position = 0))
axis(1, at = seq(0, 200, 15),labels = seq(0, 200, 15), cex.axis = 1, cex.lab = 0.7)
par(mgp = c(axis.title.position = 1, axis.label.position = 0, axis.line.position = -1))
axis(2, at = seq(0, 1, 0.2), las = 1, cex.axis = 1, labels = seq(0, 1, 0.2), cex.lab = 0.7)
mtext("Cumulative Germination Probability", side = 2, line = 2, cex = 1)
mtext("Time in days", side = 1, line = 2, cex = 1)
# Add the estimate of time to event (germination) using Kaplan-Meier method 

## Estimate of survival curve
my.fit12 <- survfit(Surv(G_Day, Germination) ~ Treatment, data = g12)
## use fun = 'event' in lines function to draw the event (1-survival probability)
lines(my.fit12, fun = 'event', col = "black", lty = 3:2, lwd = 2)
legend(70, 1, lty = 3:2, box.lty = 0, cex = 1, lwd = 1.2,
       legend = c("Hand-self pollination", "Hand-cross pollination"))

# Add the cure model lines to the graph
## Save the probabilities estimated (1-est) by the Cure Model in a dataframe
S12 <- summary(Surv_Cure12, t = seq(from = 0, to = 122, by = 5), type = "survival", tidy = T)
NewS12 <- as.data.frame(S12 %>% mutate(EST = 1 - est))

## Filter the estimates by year
S_HSP12 <- NewS12 %>% filter(Treatment == "HSP")
S_HCP12 <- NewS12 %>% filter(Treatment == "HCP")

## Add lines to the graph
lines(S_HSP12$time, S_HSP12$EST, type = "l")
lines(S_HCP12$time, S_HCP12$EST, type = "l")
```



```{r echo=TRUE, message=FALSE, warning=FALSE}
ACUM_13 <- ACUM %>% dplyr::filter(Year == "2013")
plot(ACUM_13$Day, ACUM_13$EVENT, 
     pch = c(19, 1)[as.numeric(as.factor(ACUM_13$Treatment))], 
     cex = 2, 
     axes = F,
     xlim = c(0, 185), 
     ylim = c(0, 1), ylab = "", xlab = "")
par(mgp = c(axis.title.position = 0, axis.label.position = 0.9, axis.line.position = 0))
axis(1,at = seq(0, 185, 10),labels = seq(0, 185, 10), cex.axis = 1, cex.lab = 0.7)
par(mgp = c(axis.title.position = 0.1, axis.label.position = 0, axis.line.position = -1))
axis(2, at = seq(0, 1, 0.2), las = 1, cex.axis = 1, labels = seq(0, 1, 0.2), cex.lab = 0.7)
mtext("Cumulative Germination Probability", side = 2, line = 2, cex = 1)
mtext("Time in days", side = 1, line = 2, cex = 1)

# Add the estimate of time to event (germination) using Kaplan-Meier method
## Estimate of survival curve
my.fit13 <- survfit(Surv(G_Day, Germination) ~ Treatment, data = g13)
## use fun = 'event' in lines function to draw the event (1-survival probability)
lines(my.fit13, fun = 'event', col = "black", lty = 3:2, lwd = 2)
legend(0.5, 0.97, lty = 3:2, box.lty = 0, cex = 0.8, lwd = 1.2,
       legend = c("Hand-self pollination", "Hand-cross pollination"))

# Add the cure model lines to the graph
## Save the probabilities estimated (1-est) by the Cure Model in a dataframe

S13 <- summary(Surv_Cure13, t = seq(from = 0, to = 185, by = 5), type = "survival", tidy = T)
NewS13 <- as.data.frame(S13 %>% mutate(EST = 1-est))
## Filter the estimates by year
S_HSP13 <- NewS13 %>% filter(Treatment == "HSP")
S_HCP13 <- NewS13 %>% filter(Treatment == "HCP")
## Add lines to the graph
lines(S_HSP13$time, S_HSP13$EST, type = "l")
lines(S_HCP13$time, S_HCP13$EST, type = "l")
```




