---
title: "Survival Analysis"
author: "Ruth Delgado-Dávila"
output: html_document
---

```{r warning=FALSE, include=FALSE}
library(tidyverse)
library(survival) # Surv function
library(flexsurvcure) # flexsurvcure function
library(cmprsk) # cuminc function
library(survminer) # ggcompetingrisks function
```

# Data

```{r echo=TRUE, message=FALSE, warning=FALSE}
germination <- read.csv("https://zenodo.org/record/5091713/files/Germination_ID.csv?download=1")
head(germination)
```

# Plotting cumulative incidence by treatment 

```{r}
# Calculate the cumulative germination between treatments
g12 <- filter(germination, Year == "2012")
cum_g12 <- cuminc(
  g12$G_Day, 
  g12$Germination, 
  g12$Treatment,
  cencode = 2
  )
ggcompetingrisks(
  fit = cum_g12,
  multiple_panels = FALSE,
  xlab = "Days",
  ylab = "Cumulative seed germination",
  ylim = c(0, 1)
  )
g13 <- filter(germination, Year == "2013")
cum_g13 <- cuminc(
  g13$G_Day, 
  g13$Germination, 
  g13$Treatment,
  cencode = 2
  )
ggcompetingrisks(
  fit = cum_g13,
  multiple_panels = FALSE,
  xlab = "Days",
  ylab = "Cumulative seed germination",
  ylim = c(0, 1))
```

# Cure model
A parametric mixture cure model is applied to compare probability survival curves per treatment by year.
The analysis required the creation of survival objects, which are created using the surv function. A survival object consists of two columns:    
- First column is the  time or censored time.\medskip
- Second column is the censoring indicator, indicating right censored data.  

```{r, echo=TRUE, message=FALSE, warning=FALSE}
surv_cure12 <- flexsurvcure(
  Surv(G_Day, Germination) ~ Treatment, 
  data = g12, link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
surv_cure12_null <- flexsurvcure(
  Surv(G_Day, Germination) ~ 1, 
  data = g12,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
mp_norm1 <- surv_cure12$loglik
mp_norm0 <- surv_cure12_null$loglik
chisq <- 2 * (mp_norm1 - mp_norm0)  
chisq
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
surv_cure13 <- flexsurvcure(
  Surv(G_Day, Germination) ~ Treatment, 
  data = g13,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
surv_cure13_null <- flexsurvcure(
  Surv(G_Day, Germination) ~ 1, 
  data = g13,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
mp_norm1 <- surv_cure13$loglik
mp_norm0 <- surv_cure13_null$loglik
chisq <- 2 * (mp_norm1 - mp_norm0)  
chisq
```




