---
title: "Survival Analysis"
author: "Ruth Delgado-Dávila"
output: html_document
---

```{r warning=FALSE, include=FALSE}
library(tidyverse)
library(survival) # Surv function
library(flexsurvcure) # flexsurvcure function
library(cmprsk) # cuminc function
library(survminer) # ggcompetingrisks function
```

This document show how to analyze the **time to seed germination** using **Survival Analysis**. The data used here correspond to seeds of 38 plants of *Ipomoea hederacea* (Convolvulaceae) obtained after the aplication of the following experiments: 

* Hand-self pollination: flowers were manually pollinated with pollen from the two longest anthers of the flower.\medskip
* Hand-cross pollination: emasculated flowers were manually pollinated with pollen obtained from the two longest anthers of two-three donor plants.  

# Data
```{r echo=TRUE}
germination <- read.table("../data/germination.csv", header = TRUE, sep = ",")
head(germination)
```

# Plotting cumulative incidence by treatment 

```{r}
# Calculate the cumulative germination between treatments
g12 <- filter(germination, year == "2012")
cum_g12 <- cuminc(
  g12$day, 
  g12$germination, 
  g12$treatment,
  cencode = 2
  )
ggcompetingrisks(
  fit = cum_g12,
  multiple_panels = FALSE,
  xlab = "Days",
  ylab = "Cumulative seed germination",
  ylim = c(0, 1)
  )
g13 <- filter(germination, year == "2013")
cum_g13 <- cuminc(
  g13$day, 
  g13$germination, 
  g13$treatment,
  cencode = 2
  )
ggcompetingrisks(
  fit = cum_g13,
  multiple_panels = FALSE,
  xlab = "Days",
  ylab = "Cumulative seed germination",
  ylim = c(0, 1))
```

# Cure model
A parametric mixture cure model is applied to compare probability survival curves per treatment by year.
The analysis required the creation of survival objects, which are created using the surv function. A survival object consists of two columns:    
- First column is the  time or censored time.\medskip
- Second column is the censoring indicator, indicating right censored data.  

```{r, echo=TRUE, message=FALSE, warning=FALSE}
surv_cure12 <- flexsurvcure(
  Surv(day, germination) ~ treatment, 
  data = g12, link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
surv_cure12_null <- flexsurvcure(
  Surv(day, germination) ~ 1, 
  data = g12,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
mp_norm1 <- surv_cure12$loglik
mp_norm0 <- surv_cure12_null$loglik
chisq <- 2 * (mp_norm1 - mp_norm0)  
chisq
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
surv_cure13 <- flexsurvcure(
  Surv(day, germination) ~ treatment, 
  data = g13,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
surv_cure13_null <- flexsurvcure(
  Surv(day, germination) ~ 1, 
  data = g13,
  link = "logistic", 
  dist = "lnorm", 
  mixture = T
  )
mp_norm1 <- surv_cure13$loglik
mp_norm0 <- surv_cure13_null$loglik
chisq <- 2 * (mp_norm1 - mp_norm0)  
chisq
```




